---
title: "seascape"
output: html_document
date: "2022-10-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r }
# remotes::install_github("marinebon/seascapeR", force = T)
# remotes::install_github("spatstat/spatstat.data")
# install.packages('seascapeR')
```
Load the library and map Seascape classes using a web map server (wms) that loads image tiles (not data) interactively (zoom, pan) from R.


##  
```{r pressure, echo=FALSE}
library(seascapeR)

# variables
sanctuary   = "mbnms"          # or see: ?get_url_ply
ss_dataset  = "global_monthly" # or "global_8day"
ss_var      = "CLASS"          # or "P"
date_beg    = "2020-01-01"
date_end    = "2021-01-01"

# paths
dir_data = here::here("data_ss")
dir_ply  = glue::glue("{dir_data}/ply")
dir_grd  = glue::glue(
  "{dir_data}/{sanctuary}_{ss_dataset}")
ts_csv   = glue::glue(
  "{dir_data}/{sanctuary}_{ss_dataset}_{ss_var}.csv")

# get sanctuary polygon
ply <- get_url_ply(
  sanctuary = sanctuary, 
  dir_ply   = dir_ply)
ply


# get SeaScape dataset information
ss_info <- get_ss_info(dataset = ss_dataset)
ss_info

# map SeaScape using web map server (wms) image tiles
map_ss_wms(ss_info, ply, ss_var = ss_var)

```


```{r}

# get SeaScape grids within polyon for date range 
grds <- get_ss_grds(
  ss_info, ply, 
  ss_var    = ss_var, 
  date_beg  = date_beg, 
  date_end  = date_end,
  dir_tif   = dir_grd)

# get first grid, a raster layer in the raster stack grds
grd <- raster::raster(grds, 1)

# map SeaScape grid
map_ss_grd(grd)

```


```{r}
# summarize SeaScape grids into a time series table
tbl <- sum_ss_grds_to_ts(grds, ts_csv = ts_csv)
tbl

# plot SeaScape time series
plot_ss_ts(tbl, show_legend = "always")

```

Interactivity
Note that when you run the code above in any of these R environments, you get an interactive visualization:

RStudio Console;
knitted html from an Rmarkdown document; or
a Shiny app.
The interactive map allows you to zoom and pan. The time series has sliders to change the time window and hover over to get values.

For more on how to use the seascapeR functions and view these interactive visualizations, see the Getting Started article.

Data files
Data files were cached by setting path variables and feeding as arguments to functions in the following order:

get_url_ply(): Based on dir_ply, save *.zip, unzip and shapefile with file components *.shp, etc, readable by any GIS program.
get_ss_grds(): Based on dir_grd, save grids (aka rasters) as GeoTIFs (*.tif) with filenames of form “grd_ ss_var _ date .tif”, readable by any GIS program.
plot_ss_ts(): Based on ts_csv, save the table as a comma-seperated value (*.csv) file, readable by any spreadsheet program.


```{r}

fs::dir_tree(dir_data)

```

https://marinebon.github.io/p2p/methods_data_science.html
https://ghsl.jrc.ec.europa.eu/download.php
https://globalfishingwatch.org/data-download/datasets/public-fishing-effort